Community Data Science meets Digital Heritage: Controversies about World Heritage Sites in the Online Community of Wikipedia

### 2019-05-11 code/paper.Rmd
page ages

@@

## Locality

@@

ggplot(randoms_revision_history_page_details_ip_prop,
       aes(prop_ip_addresses)) +
  geom_histogram() +
  theme_minimal()
```

@@

```{r}
# Histogram of proportion of anonymous edits 
# TODO and random sites
ggplot(revision_history_page_details_ip_addresses_countries,
       aes(prop_anon_of_all_edits)) +
  geom_histogram() +
  theme_minimal()
```

@@

@@

highlight.sector(south_america, track.index = 1, col = "brown", 
    text = "South America", cex = 0.8, text.col = "white", niceFacing = TRUE)
```

@@

```{r}
# get first and last edit, and inscription year
library(lubridate)
page_data_for_all_pages_rh_tbl_edit_start_end_span <- 
page_data_for_all_pages_rh_tbl %>% 
  group_by(Site) %>% 
  summarise(first_edit_date = min(rh_date),
            last_edit_date = max(rh_date),
            edit_span = last_edit_date - first_edit_date) %>% 
  mutate(Site = as.numeric(Site)) %>% 
  arrange(Site) %>% 
  bind_cols(page_data_for_all_pages) %>% 
  mutate(inscription_year = as.Date(ISOdate(Year_num, 1, 1))) %>% 
  mutate(diff_first_edit_inscription_year = as.numeric(ymd_hms(first_edit_date) - as.POSIXlt(inscription_year)) / 365 )

ggplot(page_data_for_all_pages_rh_tbl_edit_start_end_span, 
       aes(diff_first_edit_inscription_year)) +
  geom_histogram() +
  theme_minimal(base_size = 12)  +
  geom_vline(xintercept = 0, colour = "red") +
  ylim(0, 60) +
  annotate("text", x = -10, y = 55, label = "Page appeared\nbefore inscription\n←", size = 10) +
  annotate("text", x = 10,  y = 55, label = "Page appeared\nafter inscription\n→", size = 10)

@@

```{r}
randoms_revision_history_page_details_ip_location %>% 
  group_by(page) %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(page, the_m) %>% 
  tally() %>% 
  filter(n() > 130) %>%   # explore this value
  ggplot() +
  geom_line(aes(the_m,
             n)) +
  facet_wrap( ~ page, scales = "free")
```

@@

## Technicity

bots

What do bots do?

Do people undo bot actions often?

>>>>

## Some sort of clustering

```{r}
normalize = function(x) (x-min(x))/(max(x)-min(x))

#--------------------------------------------------------------------------
# take a look at the distribution of page variables
some_page_variables <- 
  page_data_for_all_pages %>% 
  select(page_wordcount, 
         page_wikilinks_out,
         page_wikilinks_in,
         page_cited_items_on) %>% 
  mutate(page_wikilinks_out_per_word = page_wikilinks_out / page_wordcount,
         page_cited_items_on_per_word = page_cited_items_on / page_wordcount)

# umap
library(uwot)
library("tmap")
data("World")

pages_umap_input <- 
  some_page_variables  %>% 
  mutate(page_wordcount_norm = normalize(page_wordcount),
         page_wikilinks_in_norm= normalize(page_wikilinks_in),
         page_wikilinks_out_per_word_norm = normalize(page_wikilinks_out_per_word),
         page_cited_items_on_per_word_norm = normalize(page_cited_items_on_per_word)) %>% 
  select(-page_wordcount,
         -page_cited_items_on, 
         -page_wikilinks_out,
         -page_wikilinks_in,
         -page_wikilinks_out_per_word,
         -page_cited_items_on_per_word) %>% 
  bind_cols(., page_data_for_all_pages[ , 'country'] ) %>% 
  filter_all(all_vars(!is.na(.))) %>% 
  left_join(World %>% 
              select(name, continent), 
            by = c('country' = 'name'))

# compute umap

pages_umap_input_selected <- 
  pages_umap_input %>% 
  select(-country, 
         -continent,
         -geometry
         )

pages_umap_output <- 
  pages_umap_input_selected %>% 
  umap(., 
       n_neighbors = 60, 
       min_dist = 0.7,
       nn_method = "annoy",
       init = "spca") %>% 
  as_tibble()

# compute hdbscan clusters
library(dbscan)
hdbscan_out <- hdbscan(pages_umap_output, 
                       minPts = 5)

table(hdbscan_out$cluster)

main_plot <- 
ggplot(pages_umap_output,
       aes(V1, V2)) +
  geom_point(
    aes(colour = factor(hdbscan_out$cluster),
        size =  pages_umap_input_selected$page_wordcount_norm  ,
        shape = pages_umap_input$continent
        )) +
  scale_size(range = c(2,7), guide = FALSE) +
  scale_shape(name = "Continent") +
  scale_color_viridis_d(guide = FALSE) +
  theme_minimal() +
  theme(legend.position = c(0.1, 0.8)) +
  xlab("") +
  ylab("")

# train a feature-selecting classificator like random forests on 
# the cluster labels

rand_forest_input <- 
  pages_umap_input_selected %>% 
  mutate(clus = hdbscan_out$cluster) %>% 
  filter(clus != 0)

library(caret)

fit <- train(
  clus ~ .,
  data = rand_forest_input,
  method = "ranger",
  trControl = trainControl(method="cv", 
                            number = 10, 
                            allowParallel = TRUE, 
                            verbose = TRUE),
  importance = 'permutation')

fit
var_imp_tbl <- tibble(var = row.names(varImp(fit)$importance),
                      imp = varImp(fit)$importance$Overall)

sub_plot <- 
ggplot(var_imp_tbl,
       aes(reorder( var, -imp ),
           imp)) +
  geom_col() +
  coord_flip() +
  xlab("") +
  ylab("") +
  theme_nogrid(base_size = 6)

# plot plus subplot
main_plot + 
  annotation_custom(ggplotGrob(sub_plot), 
                    xmin = 3, 
                    xmax = 6, 
                    ymin= 6.5, 
                    ymax= 3.5)

ggsave("figures/wh_country_umap_hdbscan.png")
```

@@

## Spatial patterns

@@

both_ip_props <- 
bind_rows(
revision_history_page_details_ip_addresses_countries %>% 
  ungroup %>% 
  select(prop_anon_of_all_edits) %>% 
  mutate(source = "World Heritage Sites", 
         prop_ip_addresses = prop_anon_of_all_edits),

randoms_revision_history_page_details_ip_prop %>% 
    ungroup %>% 
  select(prop_ip_addresses) %>% 
    mutate(source = "Random pages"))

ggplot(both_ip_props,
       aes(fill = source, 
           prop_ip_addresses)) +
  geom_density(alpha = 0.3) +
   scale_fill_viridis_d(labels = c("Random", 
                               "World Heritage")) +
  theme_minimal() +
  xlab("Proportion of all edits that are anonymous")

ggsave("figures/proportion_of_IP_edits.png")
```

@@

```{r}
# Histogram of proportion of anonymous edits 
# TODO and random sites
ggplot(revision_history_page_details_ip_addresses_countries,
       aes(prop_anon_of_all_edits)) +
  geom_histogram() +
  theme_minimal(base_size = 12) +
  xlab("Proportion of all edits that are anonymous")
```

@@

# stuff that doesn't work

@@

@@

```{r}
# get first and last edit, and inscription year
library(lubridate)
page_data_for_all_pages_rh_tbl_edit_start_end_span <- 
page_data_for_all_pages_rh_tbl %>% 
  group_by(Site) %>% 
  summarise(first_edit_date = min(rh_date),
            last_edit_date = max(rh_date),
            edit_span = last_edit_date - first_edit_date) %>% 
  mutate(Site = as.numeric(Site)) %>% 
  arrange(Site) %>% 
  bind_cols(page_data_for_all_pages) %>% 
  mutate(inscription_year = as.Date(ISOdate(Year_num, 1, 1))) %>% 
  mutate(diff_first_edit_inscription_year = as.numeric( as.POSIXlt(inscription_year) - ymd_hms(first_edit_date) ) / 365 )

page_data_for_random_rh_tbl_edit_start_end_span <- 
randoms_revision_history_page_details %>% 
  group_by(page) %>% 
  summarise(first_edit_date = min(rh_date),
            last_edit_date = max(rh_date),
            edit_span = last_edit_date - first_edit_date) %>% 
  mutate(page = as.numeric(page)) %>% 
  arrange(page)

bind_rows(

  page_data_for_all_pages_rh_tbl_edit_start_end_span %>% 
    select(edit_span) %>% 
    mutate(source = "World Heritage"),

  page_data_for_random_rh_tbl_edit_start_end_span %>% 
     select(edit_span) %>% 
    mutate(source = "Random")

) %>% 
  ggplot() +
geom_density(aes(as.numeric(edit_span), 
                 fill = source),
                 alpha = 0.3) +
  # scale_x_log10() +
  scale_fill_viridis_d() +  
  theme_minimal(base_size = 12) + 
  xlab("Age of Wikipedia page (days)")

ggsave("figures/wh_wikpedia_pages_ages.png")

Time between inscription year and first edit

ggplot(page_data_for_all_pages_rh_tbl_edit_start_end_span,
       aes(Year_num, 
           diff_first_edit_inscription_year)) +
  geom_jitter(alpha =  0.1, size = 3)  +
  geom_vline(xintercept = 2001, colour = "red") +
  geom_hline(yintercept = 0, colour = "red") +
  theme_minimal(base_size = 12) + 
  xlab("Year of inscription on World Heritage list") +
  ylab("Difference between year of\ninscriptionand first edit on Wikipedia")

ggplot(page_data_for_all_pages_rh_tbl_edit_start_end_span,
       aes(Year_num, 
           first_edit_date)) +
  geom_jitter(alpha =  0.1, size = 3) +
  theme_minimal(base_size = 12) + 
  xlab("Year of inscription on World Heritage list") +
  ylab("first edit on Wikipedia")

ggplot(page_data_for_all_pages_rh_tbl_edit_start_end_span,
       aes(Year_num)) +
  geom_histogram() + 
  theme_minimal(base_size = 12) +
  scale_x_continuous(breaks = seq(1975, 2019, 1)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  xlab("Year of inscription on World Heritage list")

ggsave("figures/wh_wikpedia_pages_inscription_year_hist.png")

ggplot(page_data_for_all_pages_rh_tbl_edit_start_end_span, 
       aes(diff_first_edit_inscription_year)) +
  geom_histogram() +
  theme_minimal(base_size = 12)  +
  geom_segment(x = 0, xend = 0, y = 0, yend = 50, colour = "red") +
   geom_segment(x = -10, xend = -10, y = 0, yend = 50, colour = "blue") +
   geom_segment(x = -22, xend = -22, y = 0, yend = 50, colour = "blue") +
  ylim(0, 70) +
  annotate("text", x = 10, y = 59, label = "Page appeared\nbefore inscription\n→", size =8) +
  annotate("text", x =-10,  y = 59, label = "Page appeared\nafter inscription\n←", size = 8) +
  annotate("text", x =-22,  y = 5, label = "Inscribed in 1983→", 
           size = 3, colour = "white", hjust = 1  ) +
  annotate("text", x =-10,  y = 5, label = "Inscribed in 1994→", 
           size = 3, colour = "white", hjust = 1 ) +
  annotate("text", x =0,  y = 5, label = "Inscribed in 2001→", 
           size = 3, colour = "white", hjust = 1 ) +
  xlab("Difference between year of inscription on World Heritage list and first edit on Wikipedia")

ggsave("figures/wh_wikpedia_pages_diff_time_inscription.png")

@@

```{r}
randoms_revision_history_page_details_ip_location %>% 
  group_by(page) %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(page, the_m) %>% 
  tally() %>% 
  filter(n() > 130) %>%   # explore this value
  ggplot() +
  geom_line(aes(the_m,
             n)) +
  facet_wrap( ~ page, scales = "free")

@@
